Light inextensible strings under tension in the Schwarzschild geometry

### 2019-03-21 string_ajp.tex
slight rearrangement of diagrams plus some typos

Here I solve equation~\ref{rdashdash} using the \verb+deSolve+ package
for ordinary differential equations~\cite{soetart2010}.  Solutions
fall into two distinct classes: those with $r>1$ at all points, and
those which attain $r=1$, in which case the string is tangential to
the event horizon, unless it is radial.

Taking strings for which $r>1$ first,
Figure~\ref{closest_approach_non_self_intersecting} shows a sequence
of non-self-intersecting numerical solutions to
equation~\ref{rdashdash}, rotated so that the tangential point of
closest approach occurs at~$\phi=0$.
Figure~\ref{closest_approach_self_intersecting} shows the
corresponding diagram for self-interesecting strings
(Figure~\ref{light_closest_approach} shows comparable null geodesic
diagram for comparison).  A bisection technique reveals that if the
closest approach is greater than a critical value of~$\simeq
1.076109317(1)$, the string does not self-intersect; if the closest
approach is less than this value, then the string crosses itself
at~$\phi=\pi/2$.  As discussed above, if $\left(r'\right)^2=2r(r-1)$,
the string has a point of inflection at which point the radius of
curvature becomes infinite.  Figure~\ref{curvature_switch} shows
non-self intersecting strings rotated so that the point of inflection
occurs at~$\phi=0$; we see that each string has a finite length in
which it is convex to the origin and two semi-infinite stretches in
which the curvature is negative.

The cases where the string is tangent to the event horizon
at~$\phi=0$, including a radial string, are shown in
Figures~\ref{one_free_end_r_equals_2} and~\ref{fixed_EH_intersection}
for strings tangential to or missing the Schwarzschild radius
respectively.  There are other ways to parametrize the strings; Figure
\ref{strings_r_equals_2}, for example, shows strings passing
through~$(2,0)$; Figure~\ref{light_r_equals_2} shows light paths for
comparison.

>>>>

Here I solve equation~\ref{rdashdash} using the \verb+deSolve+ package
for ordinary differential equations~\cite{soetart2010}.  Non-radial
solutions fall into two distinct classes: those with $r>1$ at all
points, and those which attain $r=1$, in which case the string is
tangential to the event horizon.

We first consider strings for which $r>1$ everywhere (that is,
completely exterior strings).
Figure~\ref{closest_approach_non_self_intersecting} shows a sequence
of non-self-intersecting numerical solutions to
equation~\ref{rdashdash}, rotated so that the tangential point of
closest approach occurs at~$\phi=0$.
Figure~\ref{closest_approach_self_intersecting} shows the
corresponding diagram for self-interesecting strings
(Figure~\ref{light_closest_approach} shows equivalent null geodesics
for comparison).  A bisection technique reveals that if the closest
approach is greater than a critical value of~$\simeq 1.076109317(1)$,
the string does not self-intersect; if the closest approach is less
than this value, then the string crosses itself at~$\phi=\pi/2$.  As
discussed above, if $\left(r'\right)^2=2r(r-1)$, the string has a
point of inflection at which point the radius of curvature becomes
infinite.  Figure~\ref{curvature_switch} shows non-self intersecting
strings rotated so that the point of inflection occurs at~$\phi=0$; we
see that each string has a finite length in which it is convex to the
origin and two semi-infinite stretches in which the curvature is
negative.

There are other ways to parametrize completely exterior strings;
Figure \ref{strings_r_equals_2}, for example, shows strings passing
through~$(2,0)$.  Figure~\ref{light_r_equals_2} shows light paths for
comparison.

The cases where the string attains $r=1$ are tangent to the event
horizon (or, exceptionally, purely radial).
Figure~\ref{fixed_EH_intersection} shows strings tangential to the
event horizon at~$\phi=0$ and Figure~\ref{one_free_end_r_equals_2}
shows tangential strings passing through $(2,0)$.
